---
title: "Texas CPS Data and County Graphics"
author: "Ben Jacobson"
date: '2017-11-25'
categories: Draft
nocite: |
  @R-RSocrata
  @R-tidyverse
  @R-tigris
  @R-broom
  @R-hidden
slug: texas-cps-data-and-county-graphics
tags:
- Socrata
- Texas Open Data
- tigris
bibliography: packages2.bib
---
```{r include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, include = FALSE}
library(hidden)
library(tidyverse)
library(RSocrata)
library(tigris)
library(broom)
knitr::write_bib(.packages(), "packages2.bib") 
```

This exercise will begin by exploring confirmed CPS abuse/neglect allegations in Texas from 2008 to 2016. The data is publicly available on the open data portal of [Texas.gov](https://data.texas.gov/), specifically, [CPS Abuse/Neglect Investigations Types of Abuse - Confirmed Allegations by County and Region FY08- FY16](https://data.texas.gov/Social-Services/CPS-Abuse-Neglect-Investigations-Types-of-Abuse-Co/9qeb-ewj9).  

Get my Socrata token...
```{r}
tok <- hidden::get_Texas_Socrata_token()
```

Get dataset... 
```{r}
dsLocation <- "https://data.texas.gov/resource/3pwr-c7hg.csv"
df <- read.socrata(dsLocation, app_token = tok)
```

General explorations and plotting...
```{r}
summary(df)
```

```{r}
unique(df$type_of_allegation)
```

```{r}
df$type_of_allegation <- as.factor(df$type_of_allegation)
df$fiscal_year <- lubridate::make_date(df$fiscal_year)
```

```{r plot1}
df %>%
    select(fiscal_year, type_of_allegation, confirmed_allegations) %>%
    group_by(fiscal_year, type_of_allegation) %>%
    summarize(confirmed_allegations = sum(confirmed_allegations, na.rm=TRUE)) %>%
    ggplot(aes(x=fiscal_year, y=confirmed_allegations)) +
    geom_point() + 
    facet_wrap(~type_of_allegation, scales = "free_y") +
    labs(title="Texas Confirmed Abuse/Neglect Cases",
         caption="Source: Texas Open Data, data.texas.gov")
    
```



Inspired by this blog post by Michael Lee on his [blog](http://www.mikelee.co/posts/), I'm going to use the tigris package to obtain a shapefile of Texas counties.

```{r, plot2}
tx <- counties("texas", cb = TRUE)
plot(tx)
```


But, how do I merge the data into this shapefile to produce the map plot that I'm seeking?

(http://zevross.com/blog/2015/10/14/manipulating-and-mapping-us-census-data-in-r-using-the-acs-tigris-and-leaflet-packages-3/#plotting-with-ggplot2)

The R help for fortify informs users "Rather than using this function, I now recommend using the broom package, which implements a much wider range of methods. fortify may be deprecated in the future."

```{r}
dfsp <- tidy(tx)
head(dfsp)
```

That outcome doesn't appear to help me join the CPS data with this spatial data frame (SDF). I'm also very curious if it's possible to tidyverse this workflow. It doesn't appear so, but I intend to research this. The tigris package includes a function designed for merging data to an SDF, geo_join. I wonder if I can join these on Texas county names.
```{r}
dfsp2 <- geo_join(tx, df, 'NAME', 'county')
summary(dfsp2@data)
```

This performed a left join
```{r}
glimpse(dfsp2@data)
```

We have a single NA in fiscal_year that I'm interested in, but it does look like it's appropriate.
```{r}
dfsp2@data[is.na(dfsp2@data$fiscal_year),]
```

# References  
